---
title: "Case Estimates"
output:
    github_document:
        toc: true
        toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(getwd())
```

## Description
This script estimates the number of cases in different districts of Cameroon or any other countries based on epidemiological data, population data, and rural-urban data. 
It first loads the necessary libraries and data, then reclassifies the rural-urban data to exclude urban areas.  
It then calculates the number of cases by multiplying the population data (excluding urban areas) by the epidemiological map. The cases are then aggregated by district, and the results are plotted on a map. 

We need to make sure that the rural urban map and population count map should match in terms of resolution and extent. They also need to be from the closest time period possible with the epidemiological map to get the most accurate estimates.

## Load libraries

```{r, message=FALSE, warning=FALSE}
library(raster)
library(sf)
library(exactextractr)
library(geodata)
library(ggplot2)
library(RColorBrewer)
library(ggspatial)
```

## Set the output directory for downloaded data

```{r}
output_dir <- getwd()
country_boundary <- gadm("CMR", level = 0, path = output_dir) 
country_boundary_sf <- country_boundary %>% st_as_sf()
country_boundary_3 <- gadm("CMR", level = 3, path = output_dir) 
country_boundary_3_sf <- country_boundary_3 %>% st_as_sf()
```

## Load the epi map, population data, and rural-urban data

```{r}
epi_map <- raster("data/day_temp_cameroon_2010_12_1km.tif") # dummy data as example, need to input prevalence map here
pop_count <- raster("data/Pop_count_cameroon_2010_12_1km.tif")
rur_urb_data <- raster("data/2024-01-26_rural_urban_Cameroon_2010_1km.tif")
```

## Sanity check

```{r}
par(mfrow = c(2, 2))
plot(epi_map, main = "Epi Map")
plot(pop_count, main = "Population Data")
plot(rur_urb_data, main = "Rural-Urban Data")
plot(country_boundary, main = "Country Boundary")
par(mfrow = c(1, 1))
```

## Exclude urban areas

```{r}
# Reclassify the rural-urban data raster (rur_urb_data) into a new raster (rururb_nourban).
# In the new raster, values less than 1 and values 20 or greater are set to NA (representing negative (invalid) areas and urban areas, respectively),
# and all other values are set to 1 (representing rural and non-urban areas).

rururb_nourban <- reclassify(rur_urb_data, c(-Inf, 1, NA,
                    1, 10, 1,
                    10, 20, 1,
                    20, Inf, NA))
plot(rururb_nourban)
```

## Get population data without urban areas

```{r}
pop_count_rururb_nourban <- pop_count * rururb_nourban

## number of cases - assuming epimap is prevalence in percent
n_cases <- pop_count_rururb_nourban * epi_map
```

## Calculate mean cases by district

```{r}
cases <- exactextractr::exact_extract(n_cases, country_boundary_3_sf, fun = "sum")
country_boundary_3_sf$cases <- cases/1000
```

## Plot aggregated cases

```{r}
mypalette <- colorRampPalette((brewer.pal(9, "Reds")))
agg_cases <- ggplot()+
  geom_sf(data = country_boundary_3_sf, aes(fill = cases), col = "black", lwd = .3) +
  geom_sf(data = country_boundary_sf, col = "black", lwd = .8, alpha = 0) + 
  scale_fill_gradientn(colours = mypalette(100)) +
  labs(fill = "Estimated aggregated cases (x1000)") +
  theme_void() +
  theme(legend.position = "bottom", legend.key.size = unit(2, 'cm'), #change legend key size
        legend.key.height = unit(.5, 'cm'), #change legend key height
        legend.key.width = unit(1.5, 'cm')) +
  guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5)) +  annotation_scale(location = "br", width_hint = 0.5)  +
  annotation_north_arrow(location = "tl", which_north = "true",
        pad_x = unit(0.15, "in"), pad_y = unit(0.25, "in"),
        style = north_arrow_fancy_orienteering)
agg_cases
```

## Session info

```{r}
sessionInfo()
```